---
title: "Transportation ROI Analysis: OKI Region"
author: "Benedict Leonardi"
---

This analysis has been integrated into the [Sprawl Paradox](https://bleonardi.github.io/True_Urban_Ranking_Project/fiscal_roi.html) project.

[View the full analysis →](https://bleonardi.github.io/True_Urban_Ranking_Project/fiscal_roi.html)

---

## Original Analysis

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(scales)

results <- read_csv("data/oki_roi_results.csv", show_col_types = FALSE) %>%
  mutate(GEOID = str_pad(as.character(GEOID), 12, pad = "0"))
summary <- read_csv("data/oki_roi_summary.csv", show_col_types = FALSE)
tracts <- st_read("data/oki_tracts.geojson", quiet = TRUE) %>%
  mutate(GEOID = str_pad(as.character(GEOID), 12, pad = "0"))

map_data <- tracts %>% inner_join(results, by = "GEOID")
```

### Key Findings

### Fiscal Productivity by Density Class

```{r summary-table, echo=FALSE}
summary %>%
  knitr::kable(digits = 2, caption = "ROI Summary by Density (Per Acre Metrics)")
```

### Spatial Distribution

```{r map, echo=FALSE}
pal <- colorNumeric(palette = "YlOrRd", domain = map_data$Revenue_Per_Acre)

leaflet(map_data) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    fillColor = ~pal(Revenue_Per_Acre),
    weight = 1, opacity = 1, color = "white", dashArray = "3", fillOpacity = 0.7,
    label = ~paste0(NAME, ": ", dollar(Revenue_Per_Acre), "/acre")
  ) %>%
  addLegend(pal = pal, values = ~Revenue_Per_Acre, opacity = 0.7,
            title = "Tax Revenue / Acre", position = "bottomright")
```

### Methodology

1. **Revenue Proxy:** Estimated using Median Home Value × Housing Units × 1.25% Tax Rate.
2. **Investment:** Federal DOT awards (USAspending) mapped to road segments (Overture Maps) and aggregated by Census Block Group.
3. **Maintenance Liability:** Estimated at $10,000 per lane-mile annually based on total road mileage per block group.
