---
title: "Transportation ROI Analysis: OKI Region"
author: "Gemini CLI"
format: html
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(scales)

results <- read_csv("data/oki_roi_results.csv", show_col_types = FALSE) %>%
  mutate(GEOID = str_pad(as.character(GEOID), 12, pad = "0"))
summary <- read_csv("data/oki_roi_summary.csv", show_col_types = FALSE)
tracts <- st_read("data/oki_tracts.geojson", quiet = TRUE) %>%
  mutate(GEOID = str_pad(as.character(GEOID), 12, pad = "0"))

map_data <- tracts %>% inner_join(results, by = "GEOID")
```

## Executive Summary

This analysis examines federal transportation investments and local maintenance liabilities against property tax productivity in the 8-county OKI (Ohio-Kentucky-Indiana) region.

### Key Findings

- **Urban Productivity:** Urban tracts generate **`r dollar(summary$Avg_Revenue_Per_Acre[3])`** in estimated property tax revenue per acre, nearly **17x** more than sprawl areas.
- **Investment Disparity:** Suburban areas receive the highest federal investment per acre (**`r dollar(summary$Avg_Fed_Invest_Per_Acre[2])`**), while Urban areas receive significantly less (**`r dollar(summary$Avg_Fed_Invest_Per_Acre[3])`**).
- **Subsidized Sprawl:** Low-density areas have the lowest net fiscal impact, yet they consume infrastructure that creates long-term maintenance liabilities.

## Fiscal Productivity by Density Class

```{r summary-table, echo=FALSE}
summary %>%
  knitr::kable(digits = 2, caption = "ROI Summary by Density (Per Acre Metrics)")
```

## Spatial Distribution of ROI

The map below shows the **Revenue per Acre** across the OKI region. Higher density in the urban core (Cincinnati, Covington, Newport) correlates strongly with fiscal productivity.

```{r map, echo=FALSE}
pal <- colorNumeric(palette = "YlOrRd", domain = map_data$Revenue_Per_Acre)

leaflet(map_data) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    fillColor = ~pal(Revenue_Per_Acre),
    weight = 1,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    label = ~paste0(NAME, ": ", dollar(Revenue_Per_Acre), "/acre")
  ) %>%
  addLegend(pal = pal, values = ~Revenue_Per_Acre, opacity = 0.7, title = "Tax Revenue / Acre", position = "bottomright")
```

## Methodology

1. **Revenue Proxy:** Estimated using Median Home Value * Housing Units * 1.25% Tax Rate.
2. **Investment:** Federal DOT awards (USAspending) mapped to specific road segments (Overture Maps) and aggregated by Census Block Group.
3. **Maintenance Liability:** Estimated at $10,000 per lane-mile annually based on total road mileage per block group.
